Algebraic File Synchronization: Adequacy and Completeness

### 2018-07-18 p2/multidir.tex
Text review

The two assumptions we used to increase the symmetry of our model were
that there is no move command,
and that
there is only one directory value, that is, directories are not differentiated
by the meta-information they contain.
In this section we describe reasons why the latter assumption may not limit the applicability
of even the current model, and we also describe encoding and decoding steps
to overcome the two restrictions.

@@

Despite the above, there can be applications where directory metadata
is considered important.
Synchronizers based on our model can be readily extended to handle them
by duplicating $\setn$. 
Given the input filesystems over $\setn$, we add a special node as an extra child under each original node,
and we encode directory metadata in a file under each directory.
It is easy to see that update detection and conflict resolution can continue as expected,
with the only exception of a potential conflict detected on one of these special nodes.
In these cases the synchronizer may prescribe creating a directory without creating
the special metadata file, which is clearly not possible to do on the target filesystem,
as creating a directory entails specifying its metadata as well.
In these cases, the synchronizer can either fall back to default values and flag the issue
for review later, or if, as suggested above, conflicts on the metadata (e.g. readable, writable and executable flags)
can be easily automated, then it could form part of the implementation.

Reintroducing the \emph{move} command can happen in a similar fashion.
Its main advantages include that 
\emph{move} commands are easier to review,
and if the synchronizer suggests or performs a move instead
of deleting a file and recreating it somewhere else, the user
can be assured that no information is lost.
During pre-processing, if an update detector is used,
\emph{move} commands still do not need to be part of the command sequences
the synchronizer operates on.
If the command sequences are derived from filesystem journals, they
can be broken into two updates to allow the existing reconciliation
algorithm to operate.

>>>>

The two assumptions we used to increase the symmetry of our model were
that there is no \emph{move} command,
and that
there is only one directory value, that is, directories are not differentiated
by the meta-information they contain.
In this section we describe reasons why the latter assumption may not limit the applicability
of even the current model, and we also describe encoding and decoding steps
to overcome the two restrictions.

@@

Despite the above, there can be applications where directory metadata
is considered important.
Synchronizers based on our model can be readily extended to handle them
by duplicating $\setn$. 
Given the input filesystems over $\setn$, we add a special node as an extra child under each original node,
and we encode directory metadata in a file under each directory.
It is easy to see that update detection and conflict resolution can continue as expected,
with the only exception of a potential conflict detected on one of these special nodes.
In these cases the synchronizer may prescribe creating a directory without creating
the special metadata file, which is clearly not possible to do on the target filesystem,
as creating a directory entails specifying its metadata as well.
In these cases, the synchronizer can either fall back to default values and flag the issue
for review later, or if, as suggested above, resolution of conflicts on the metadata (e.g. readable, writable and executable flags)
can be easily automated, then it could form part of the implementation.

Reintroducing the \emph{move} command can happen in a similar fashion.
Its main advantages include that 
\emph{move} commands are easier to review,
and if the synchronizer suggests or performs a move instead
of deleting a file and recreating it somewhere else, the user
can be assured that no information is lost.
During pre-processing, if an update detector is used,
\emph{move} commands still do not need to be part of the command sequences
the synchronizer operates on.
If the command sequences are derived from filesystem journals, they
can be encoded as separate delete and create operations to allow the existing reconciliation
algorithm to operate.

### 2018-07-18 p2/ordering.tex
Text review

Let $A$ be a sequence that satisfies the conditions;
we therefore know that it contains a command $\cxynv$ on $n$
and another command $\czwnnv$ on $\nn$.
We use a proof by contradiction and assume that there are no commands on $\parent(n)$ in $A$.
Next, we create a new sequence $A'\equiv A$ in which $\cxynv$ and $\czwnnv$ are next to each other.
If they are already next to each other in $A$, there is nothing to do.
Otherwise, consider the command next to $\cxynv$ in the direction where $\czwnnv$ is.
Let this command be $\cqrmv$.
If $\czwnnv$ is to the right, then $A$ looks like the following:
\[ A = \cdots\cc\cxynv\cc\cqrmv\cc\cdots\cc\czwnnv\cc\cdots \]
If $n\unrel m$, then swap $\cxynv$ and $\cqrmv$. Based on \cref{ax_separate_commute} we know that the new
sequence is equivalent to $A$.
Otherwise, we know $m\neq\parent(n)$ as there are no commands on $\parent(n)$, and so
from \cref{ax_distantrel_breaks} we get $A\equiv\cbrk$ which contradicts our assumptions.
(Note that $A$ cannot contain assertion commands.)
By repeating this step we can therefore convert $A$ into $A'$ where $\cxynv$ and $\czwnnv$ are neighboring commands.
However, then \cref{ax_distantrel_breaks} applies to the sequence and therefore $A\equiv A'\equiv\cbrk$ which
is again a contradiction.
\end{proof}

@@

\begin{proof}
First, we prove, by contradiction, that if $T$ is a simple sequence and $T\equiv S$, then $T\in\orderset{S}$.
We use an inverse proof and assume $T\equiv S$ and $T\not\in\orderset{S}$.
Then, from \cref{equiv_simple_same_commands} we know $T$ is a permutation of $S$,
and from \cref{order_is_only_possible} we know $T\equiv\cbrk$ which
is a contradiction as $T\equiv \wrksx{S}$.

Next, we prove that if $T\in\orderset{S}$, then $T\equiv S$.
We proceed by induction on the length of $S$.
Our base cases are $\emptyseq$, where $\orderset{\emptyseq}=\{\emptyseq\}$,
and one-long sequences, where this is trivially true.
In our induction step we assume that $T^*\in\orderset{S^*}\Rightarrow T^*\equiv S^*$ holds
for all sequences of length $i$ or less, where $i\geq 1$.

Let us consider $T\in\orderset{S}$ where $T$ and $S$ are of length $i+1$.
Let the first command in $S$ be $\cxynv$, that is, $S=\cxynv\cc S_0$.
We proceed by a nested induction on the number of commands before $\cxynv$ in $T$.
Our base case is when it is the first command, which means $T=\cxynv\cc T_0$.
Then from \cref{subseq_in_orderset} we know $T_0\in\orderset{S_0}$
and from the induction hypothesis $T_0\equiv S_0$ from which we get $T\equiv S$.

@@

Also, let $I^S_\indx$ be the index of the command on $n_\indx$ in $S$.
For all $\indx$, both the commands on $n_\indx$ and $n_{\indx+1}$ must be either
construction commands, or destruction commands, depending on
whether $I^S_\indx<I^S_{\indx+1}$ or $I^S_{\indx+1}<I^S_\indx$.
This is because otherwise, \cref{simple_distant_pairs} would not hold for $S$.

@@

It follows from \cref{simple_reorder_equiv} that given a simple set of commands $S$,
which can be applied to a filesystem in some order without breaking it,
this order will be in $\orderset{S}$,
and all sequences in $\orderset{S}$
represent the same partial endofunction on filesystems.
This allows us to treat such a set as a sequence for any purpose
where the internal order of the commands is irrelevant.
\begin{mydef}[$\ordp{S}$]
For a simple set or sequence of commands $\wrksx{S}$,
$\ordp{S}$ is the unique partial endofunction on filesystems
defined by any order in $\orderset{S}$.
% We leave $\ordp{S}$ undefined if such an order does not exist.
\end{mydef}

Building on the results above, we finish by proving an important result,
namely, that command sequences can be made simple using syntactical operations.

\begin{mylem}\mlabel{can_simplify}
Each sequence of commands $\wrksx{S}$
can be transformed into a simple sequence
$T\equiv S$ by applying repeatedly the following transformations:
{\rm(i)} swap neighboring commands on incomparable nodes;
{\rm(ii)} simplify neighboring commands on the same node;
{\rm(iii)} delete assertion commands.
\end{mylem}
\begin{proof}
We can remove all assertion commands, as based on
\cref{ax_assert} the resulting sequence extends the original one.
We therefore assume that $S$ contains no assertion commands.

@@

We aim to find a simple sequence that is equivalent to $S$.
If $S$ is already simple, there is nothing to do.
If it is not simple, let $\ucx$ be the first command
on a node $n$ that is already present in $S$,
and let $\ucy$ be the previous command on $n$.
Splitting $S$ around these commands we get
\[ S = S_0 \cc \ucy \cc S_1 \cc \ucx \cc S_2. \]

@@

If $n=\parentf{m'}$,
from \cref{simple_distant_pairs} 
and the relationship of nodes in construction and destruction pairs
we know that the subsequence
$\ucy\cc\ucz$ must be a construction pair,
and both $\ucy$ and $\ucz$ must be construction commands.
But we also know that $\ucz\cc\ucx$ must be a destruction pair,
and both $\ucz$ and $\ucx$ must be destruction commands.
This is a contradiction as a command cannot be both.
We arrive at the same contradiction if $m'=\parentf{n}$.

We therefore know that $S_1\indep\ucx$ and so
from \cref{ax_separate_commute}
\[ S \equiv S_0\cc\ucy\cc\ucx\cc S_1 \cc S_2. \]
We can apply \cref{ax_same_emptyseq,ax_same_singlec} to
$\ucy\cc\ucx$ and get a $S'\eqext S$ that contains one or two less commands on $n$.
Repeating the swap and the simplification we
can convert $S$ into an equivalent simple sequence.
\end{proof}

>>>>

Let $A$ be a sequence that satisfies the conditions;
we therefore know that it contains a command $\cxynv$ on $n$
and another command $\czwnnv$ on $\nn$.
By contradiction assume that there are no commands on $\parent(n)$ in $A$.
Next, we create a new sequence $A'\equiv A$ in which $\cxynv$ and $\czwnnv$ are next to each other.
If they are already next to each other in $A$, there is nothing to do.
Otherwise, consider the command next to $\cxynv$ in the direction where $\czwnnv$ is.
Let this command be $\cqrmv$.
If $\czwnnv$ is to the right, then $A$ looks like the following:
\[ A = \cdots\cc\cxynv\cc\cqrmv\cc\cdots\cc\czwnnv\cc\cdots \]
If $n\unrel m$, then swap $\cxynv$ and $\cqrmv$. Based on \cref{ax_separate_commute} we know that the new
sequence is equivalent to $A$.
Otherwise, we know $m\neq\parent(n)$ as there are no commands on $\parent(n)$, and so
from \cref{ax_distantrel_breaks} we get $A\equiv\cbrk$ which contradicts our assumptions.
(Note that $A$ does not contain assertion commands.)
By repeating this step we can therefore convert $A$ into $A'$ where $\cxynv$ and $\czwnnv$ are neighboring commands.
However, then \cref{ax_distantrel_breaks} applies to the sequence and therefore $A\equiv A'\equiv\cbrk$ which
is again a contradiction.
\end{proof}

@@

\begin{proof}
First, we prove, by contradiction, that if $T$ is a simple sequence and $T\equiv S$, then $T\in\orderset{S}$.
Assume $T\equiv S$ but $T\not\in\orderset{S}$.
Then, from \cref{equiv_simple_same_commands} we know $T$ is a permutation of $S$,
and from \cref{order_is_only_possible} we know $T\equiv\cbrk$ which
is a contradiction as $T\equiv \wrksx{S}$.

Next, we prove that if $T\in\orderset{S}$, then $T\equiv S$.
We proceed by induction on the length of $S$.
Our base cases are $\emptyseq$, when $\orderset{\emptyseq}=\{\emptyseq\}$,
and one-long sequences, when this is trivially true.
In our induction step we assume that $T^*\in\orderset{S^*}\Rightarrow T^*\equiv S^*$ holds
for all sequences of length $i$ or less, where $i\geq 1$.

Let us consider $T\in\orderset{S}$ where $T$ and $S$ are of length $i+1$.
Let the first command in $S$ be $\cxynv$, that is, $S=\cxynv\cc S_0$.
We proceed by a nested induction on $j$, the number of commands before $\cxynv$ in $T$.
Our base case is $j=0$, when it is the first command, which means $T=\cxynv\cc T_0$.
Then from \cref{subseq_in_orderset} we know $T_0\in\orderset{S_0}$
and from the induction hypothesis $T_0\equiv S_0$ from which we get $T\equiv S$.

@@

Also, let $I^S_\indx$ be the index of the command on $n_\indx$ in $S$.
By \cref{simple_distant_pairs}, for all $\indx$, both the commands on $n_\indx$ and on $n_{\indx+1}$ must be either
construction commands, or destruction commands, depending on
whether $I^S_\indx<I^S_{\indx+1}$ or $I^S_{\indx+1}<I^S_\indx$.

@@

It follows from \cref{simple_reorder_equiv} that given a simple set of commands $S$,
which can be applied to a filesystem in some order without breaking it,
this order will be in $\orderset{S}$,
and all sequences in $\orderset{S}$
represent the same partial endofunction on filesystems.
This allows us to treat such a set as a sequence for any purpose
where the internal order of the commands is irrelevant.
\begin{mydef}[$\ordp{S}$]
For a simple set or sequence of commands $\wrksx{S}$,
$\ordp{S}$ is the unique partial endofunction on filesystems
defined by any sequence in $\orderset{S}$.
% We leave $\ordp{S}$ undefined if such an order does not exist.
\end{mydef}

Building on the results above, we finish by proving an important result,
which states that command sequences can be made simple using syntactical operations only.

\begin{myth}\mlabel{can_simplify}
Each sequence of commands $\wrksx{S}$
can be transformed into a simple sequence
$S^*\eqnrv S$ by applying repeatedly the following transformations:
{\rm(i)} swap neighboring commands on incomparable nodes;
{\rm(ii)} simplify neighboring commands on the same node;
{\rm(iii)} delete assertion commands.
\end{myth}
\begin{proof}
We can remove all assertion commands, as based on
\cref{ax_assert} the resulting sequence extends the original one.
We therefore assume that $S$ contains no assertion commands.

@@

We aim to find a simple sequence that extends $S$.
If $S$ is already simple, there is nothing to do.
If it is not simple, let $\ucx$ be the first command
on a node $n$ that is already present in $S$,
and let $\ucy$ be the previous command on $n$.
Splitting $S$ around these commands we get
\[ S = S_0 \cc \ucy \cc S_1 \cc \ucx \cc S_2. \]

@@

If $n=\parentf{m'}$,
from \cref{simple_distant_pairs} 
and the relationship of nodes in construction and destruction pairs
we know that the subsequence
$\ucy\cc\ucz$ must be a construction pair,
and both $\ucy$ and $\ucz$ must be construction commands.
But we also know that $\ucz\cc\ucx$ must be a destruction pair,
and both $\ucz$ and $\ucx$ must be destruction commands.
This is a contradiction as a command cannot be both.
We arrive at the same contradiction when $m'=\parentf{n}$.

We therefore know that $S_1\indep\ucx$ and so
from \cref{ax_separate_commute}
\[ S \equiv S_0\cc\ucy\cc\ucx\cc S_1 \cc S_2. \]
We can apply \cref{ax_same_emptyseq,ax_same_singlec} to
$\ucy\cc\ucx$ and get a $S^*\eqext S$ that contains one or two less commands on $n$.
Repeating the swap and the simplification we
can convert $S$ into a simple sequence.
\end{proof}

### 2018-07-18 p2/upd-correct.tex
Minor improvements
Style tweaks
Text review

With $\ordersetsign$ we can now define our update detection algorithm.
Its inputs are the original filesystem $\FS$ which has been modified to yield $\FS^*$.
\begin{mydef}[Update detection]\mlabel{def_upddetect}~
\begin{enumerate}
    \item For each node $n$ where the value in $\FS$ and $\FS^*$ differ, add the command $\caaa{\FS(n)}{\FS^*(n)}{n}$ to
        a set of commands $U_S$. The result is a simple set of commands.
    \item Order the commands according to $\orderrel$, that is, 
        return any sequence $U$ from $\orderset{U_S}$.
\end{enumerate}
\end{mydef}

\Cref{update_works} proves that $U$ functions as expected, that is, $U\aFS=\FS^*$.
While this is trivial if we know that $U$ does not break $\FS$, we
still need to show that it is defined on $\FS$.

@@

\begin{proof}
Let us assume that some mechanism recorded all changes that occurred
to $\FS$ until it reached the state $\FS^*$.
Our set of commands is sufficient to record any change, as
there are commands for every input and output type pair, therefore
we know that the changes can be recorded as a sequence of commands $S$
where $S\aFS=\FS^*$.
($S$ will not contain assertion commands, as they do not
represent an actual change in the filesystem.)
Based on \cref{can_simplify}, there is a simple sequence $S^*$
that is equivalent to $S$.

>>>>

With the relation $\orderrel$ we can now define our update detection algorithm.
Its inputs are the original filesystem $\FS$ which has been modified to yield $\FS^*$.
\begin{mydef}[Update detection]\mlabel{def_upddetect}~
\begin{enumerate}
    \item For each node $n$ where the value in $\FS$ and $\FS^*$ differ, add the command $\caaa{\FS(n)}{\FS^*(n)}{n}$ to
        a set of commands $U^*$. The result is a simple set of commands.
    \item Order the commands according to $\orderrel$, that is, 
        return any sequence $U$ from $\orderset{U^*}$. \qedhere
\end{enumerate}
\end{mydef}

\Cref{update_works} below proves that $U$ functions as expected, that is, $U\aFS=\FS^*$.
While this is trivial if we know that $U$ does not break $\FS$, we
still need to show that it is defined on $\FS$.

@@

\begin{proof}
Let us assume that some mechanism recorded all changes that occurred
to $\FS$ until it reached the state $\FS^*$.
Our set of commands is sufficient to record any change, as
there are commands for every input and output type pair, therefore
we know that the changes can be recorded as a sequence of commands $S$
where $S\aFS=\FS^*$.
($S$ will not contain assertion commands, as they do not
represent an actual change in the filesystem.)
Based on \cref{can_simplify}, there is a simple sequence $S^*$
that extends $S$.
